---
title: "The ggplot advent calendaR - DAY 11"
output: html_notebook
---

On the eleventh day of Christmas...

...we introduced themes!

We can use themes to make finer adjustments to non-data parts of our graphic. While labs() is fine for adding labels and a title, themes allow us to choose the size, font, colour, position, etc. of that text. Today I'm going to introduce you to some complete themes.

Let's go back to our nice, clean-looking boxplots:
```{r}
trees %>%
ggplot(aes(x=type, y=height))+
  geom_boxplot(aes(fill=type), colour="black")
```

While the grey backgronud is the default in ggplot, it's certainly not a requirement. That's often one of the first things I change!

```{r}
trees %>%
ggplot(aes(x=type, y=height))+
  geom_boxplot(aes(fill=type), colour="black")+
  theme_bw()
```

Theme_bw() is a complete theme, meaning that it's a theme that can be applied as a layer that changes the look of your overall plot. Theme_gre() is the default. You can also try other complete themes, such as theme_dark(), theme_light(), theme_classic(), theme_void() and more.

```{r}
trees %>%
ggplot(aes(x=type, y=height))+
  geom_boxplot(aes(fill=type), colour="black")+
  theme_classic()
```

Personally, I like this one best!
