---
title: "The ggplot advent calendaR - DAY 8"
output: html_notebook
---

On the eight day of Christmas...

...we finally took a break from talking about colours!

Colour is a big one, but let's work with some other changes to the aesthetics of our graphic. Let's bring back the geom_point() graphic. I'm using the theme of Christmas, but of course, December has many, many holidays. Hanukkah starts on Dec. 17 this year, so let's make a Hanukkah-inspired graphic and look at different ways to change the aesthetics of our graphs.


```{r}
trees %>%
ggplot(aes(x=type, y=height))+
  geom_point(aes(), shape = 11, colour="blue", size = 6, stroke=2)
```

Here we've specified the shape, the colour, the size, and the stroke (line thickness) of the points. There are many changes we can make to aesthetics, these are just a few examples. Let's take a look at a line graph and how we can customize that.

We haven't made a line graph yet, so let's create a simple one first, before we customize it.

```{r}
trees %>%
ggplot(aes(x=height, y=xmas.magic, group=type))+
  geom_line()
```

Here we've specified that we want each line to represent a different tree type by using "group=". This goes inside aes() because it is being applied to the data (i.e., it can't be done without this specific dataset)

Now let's spruce it up. ("spruce" it up... because they're trees... hahahaha. A little Christmas cheer for you).
```{r}
trees %>%
ggplot(aes(x=height, y=xmas.magic, group=type))+
  geom_line(aes(colour=type), linetype="dashed")
```

What if we want to show the points AND the lines? Tomorrow we will add another layer to our graphic.
