---
title: "The ggplot advent calendaR - DAY 10"
output: html_notebook
---

On the tenth day of Christmas...

...we edited the text of our graphics.

Editing the look of your text and fonts in ggplot is easy, with lots of options to make it look exactly how you want it. Here's our graphic from yesterday.
```{r}
trees %>%
ggplot(aes(x=height, y=xmas.magic, group=type))+
  geom_line(aes(colour=type), linetype="dashed")+
  geom_point(aes(colour=type), size = 3, shape = 8)
```

What kind of things might we want to change? Maybe we want a title and more informative and cleaner-looking axis labels. To do this, we need to add new layer, labs() for labels. You can also use xlab(), ylab(), and ggtitle() to add them individually.

```{r}
trees %>%
ggplot(aes(x=height, y=xmas.magic, group=type))+
  geom_line(aes(colour=type), linetype="dashed")+
  geom_point(aes(colour=type), size = 3, shape = 8)+
    labs(title="Christmas Trees", x="Tree height", y="Christmas magic")
```

What if we want to make more aesthetic changes to our fonts and labels? We will need to use themes!
