---
title: "The ggplot advent calendaR - DAY 20"
output: html_notebook
---

On the twentieth day of Christmas...

... we did some position adjustments with bar plots!

First, let's create a barplot since we haven't done that yet. We'll base it on our previous sleighs subset.
```{r}
sleighs.subset %>%
  ggplot(aes(x=bells, y=reins))+
  geom_col()+
  scale_x_discrete(limits=c(4, 6, 8))+
  theme_classic()
```

The sleighs come with 4, 6, or 8 bells. So here we're displaying the counts of sleighs in each category (# of bells). But maybe we want to see some additional information in our barplot, such as the number of reins on the sleigh. I've heard that Santa takes these things super seriously, so this is completely practical and reasonable plot. Note that we have to specify that reins is a categorical variable, not a continuous one, using as.character(). In this case, we can't have 3.5 reins.
```{r}
sleighs.subset %>%
  ggplot(aes(x=bells, y=reins, fill=as.character(reins)))+
  geom_col()+
  scale_x_discrete(limits=c(4, 6, 8))+
  theme_classic()+
  scale_fill_manual(values=xmas)
```

The default is a stacked barplot (position = "stack"), but there are other ways we could display this using position adjustments. This option shows it as a percent using position = "fill".
```{r}
sleighs.subset %>%
  ggplot(aes(x=bells, y=reins, fill=as.character(reins)))+
  geom_col(position="fill")+
  scale_x_discrete(limits=c(4, 6, 8))+
  theme_classic()+
  scale_fill_manual(values=xmas)
```

We can also position them side by side using position = "dodge". Note that the red bar on the left and the green bar on the right are two bars side by side.
```{r}
sleighs.subset %>%
  ggplot(aes(x=bells, y=reins, fill=as.character(reins)))+
  geom_col(position="dodge")+
  scale_x_discrete(limits=c(4, 6, 8))+
  theme_classic()+
  scale_fill_manual(values=xmas)
```

Want a review? Try changing the name of your legend.