---
title: "The ggplot advent calendaR - DAY 21"
output: html_notebook
---

On the twenty-first day of Christmas...

...we learned about faceting!

Faceting produces smaller graphs that can be displayed alongside one another. We use facet_wrap() and facet_grid() for this. 

Let's start with facet_wrap(). Remember our line graph that looks a bit like Christmas lights? Let's use that. Here it is, as a reminder:
```{r}
trees %>%
ggplot(aes(x=height, y=xmas.magic, group=type))+
  geom_line(aes(colour=type), linetype="dashed")+
  geom_point(aes(colour=type), size = 3, shape = 8)
```

Now, instead of having all three lines on one plot, let's create three smaller plots and display them together. 
```{r}
trees %>%
ggplot(aes(x=height, y=xmas.magic))+ #when facet is defined later, no need for groups anymore. this should make the code a bit more simple. Suggesing it for the rest of the lines
  geom_line(linetype="dashed")+
  geom_point(size = 3, shape = 8)+
  facet_wrap(~type) # I think you dont need to the ncol here, as it would just understand it from the "type" categories, which are 3. if more cats maybe setting ncols would be useful
```

Let's do the same thing but using facet_grid(). The syntax is a little different, but we've produced the exact same set of plots. In our case, ".~type" puts the plots side by side.
```{r}
trees %>%
ggplot(aes(x=height, y=xmas.magic))+
 geom_line(linetype="dashed")+
  geom_point(size = 3, shape = 8)+
  facet_grid(.~type) # when looking into face_grid, I found this would also be a possibility and maybe more intuitive?
  #  facet_grid(cols = vars(type)) # I actually just found that ggplot says on facet_grid" It is most useful when you have two discrete variables, and all combinations of the variables exist in the data. if you have only one variable with many levels, try facet_wrap()" so not sure if its the most adequate here. But, I guess it can be shown but maybe explained as well in the text
```

If we want to stack our plots instead, we change up the coding within facet_grid(). In our case, "type~." stacks the plots.
```{r}
trees %>%
ggplot(aes(x=height, y=xmas.magic))+
 geom_line(linetype="dashed")+
  geom_point(size = 3, shape = 8)+
  facet_grid(rows = vars(type)) #just as I suggested above, using rows may be more intuitive than the other syntax, but you can ignore this!
```

Our datasets aren't really set up for this type of grid, but let's look at plots of reins by bells to show you how you could set up facet_grid() with multiple plots. A plot area is produced with two levels for reins and three levels for bells.
```{r}
sleighs.subset%>%
  ggplot(aes(x=reins, y=bells))+
  facet_grid(reins~bells)
```

You can also use "scales =" to adjust the scales of all or each of the plots. Let's go back to our first set of plots from today:
```{r}
trees %>%
ggplot(aes(x=height, y=xmas.magic))+
  geom_line(linetype="dashed")+
  geom_point(size = 3, shape = 8)+
  facet_wrap(~type, ncol=3)
```

To adjust the scales, we add "scales =" to facet_wrap().
```{r}
trees %>%
ggplot(aes(x=height, y=xmas.magic))+
  geom_line(linetype="dashed")+
  geom_point(size = 3, shape = 8)+
  facet_wrap(~type, scales = "free_y")
```

By choosing "free_y" we have changed the y scale from fixed to free but the x-axis remained the same. Take a look at how the y-axis scale is now different for each plot. Other options are "free_x", "fixed", or "free".

There are lots more things you can do with facets. Check out Ch. 17 of "ggplot2:Elegant Graphics for Data Analysis" for more: https://ggplot2-book.org/facet.html

Want a review? Try changing the style of the points and lines in these plots. Also, try removing the grey background!
