---
title: "The ggplot advent calendaR - DAY 19"
output: html_notebook
---

On the nineteenth day of Christmas...

... we made position adjustments!

Position adjustments are handy if you have overlapping geoms or data. You can override the default using the position argument in the geom_() functiions.

Instead of boxplots, let's look at the raw data points using a different type of geom later, geom_jitter().
```{r}
trees %>%
ggplot(aes(x=type, y=height))+
  geom_jitter(aes(colour=type))+
  theme_classic()+
  scale_colour_manual(values=xmas)

```

Position adjustments come in handy with point data like this, more so when we're working with large datasets that have many points. Let's adjust the position of our jittered points.
```{r}
trees %>%
ggplot(aes(x=type, y=height))+
  geom_jitter(aes(colour=type), position = position_jitter(width = 0.05, height = 0.5))+
  theme_classic()+
  scale_colour_manual(values=xmas)
```

Let's look at another example. We'll bring back our scatterplot of sleigh data but I'm going to cut it down a bit to make it a easier to work with. I'll do this using another tidyverse package, dplyr.

```{r}
sleighs.transposed <-t(sleighs)
sleighs.subset <-slice(sleighs, 4:12)
```

```{r}
sleighs.subset %>%
  ggplot(aes(x=deerpower, y=km_per_carrot))+
  geom_point(aes(fill=name), shape=21, size=3)+
  theme_classic()
```

That legend is fine, but let's get rid of it and instead label each point.Do you remember how to remove the legend?
```{r}
sleighs.subset %>%
  ggplot(aes(x=deerpower, y=km_per_carrot))+
  geom_point(aes(fill=name), shape=21, size=3, show.legend = F)+
  geom_text(aes(label=name))+
  theme_classic()
```

Well this might work better, but the labels are all overlapping and difficult to read. This is where position_nudge() comes in handy!

```{r}
sleighs.subset %>%
  ggplot(aes(x=deerpower, y=km_per_carrot))+
  geom_point(aes(fill=name), shape=21, size=3, show.legend = F)+
  geom_text(aes(label=name), position = position_nudge(x=10, y=0.4))+
  theme_classic()
```

And because our Stealth Sleigh is off the plot, let's fix that using what we learned on Day 13 about limits.
```{r}
sleighs.subset %>%
  ggplot(aes(x=deerpower, y=km_per_carrot))+
  geom_point(aes(fill=name), shape=21, size=3, show.legend = F)+
  geom_text(aes(label=name), position = position_nudge(x=10, y=0.4))+
  theme_classic()+
  scale_x_continuous(limits=c(0,300))
```
