---
title: "The ggplot advent calendaR - DAY 16"
output: html_notebook
---

On the sixteenth day of Christmas...

...we worked with a different type of plot and added multiple geoms to one plot!

We haven't even touched the sleigh dataset! Let's create a graphic with that so we can work with a different type of plot.
```{r}
sleighs %>%
  ggplot(aes(x=deerpower, y=km_per_carrot))+
  geom_point(aes(fill=name), shape=21, size=3)+
  theme_classic()
```

Here we've told ggplot to colour each sleigh type a different colour (some are nearly impossible to differentiate but we won't worry about that today).

Maybe we want to add a trendline to our plot. We can do this by adding a geom_smooth() layer:
```{r}
sleighs %>%
  ggplot(aes(x=deerpower, y=km_per_carrot))+
  geom_point(aes(fill=name), shape=21, size=3)+
  geom_smooth()+
  theme_classic()
```

And maybe we want to customize that line. The default is a blue line with grey background. We can change that to a black line using "colour =" and we can change the transparency of the error using "alpha =". We can also change the method of how the line is calculated using "method ="
```{r}
sleighs %>%
  ggplot(aes(x=deerpower, y=km_per_carrot))+
  geom_point(aes(fill=name), shape=21, size=3)+
  geom_smooth(colour="black", alpha=0.2, method=lm)+
  theme_classic()
```

Alpha can also be helpful when you have overlapping points.
```{r}
sleighs %>%
  ggplot(aes(x=deerpower, y=km_per_carrot))+
  geom_point(aes(fill=name), shape=21, size=3, alpha=0.5)+
  geom_smooth(colour="black", alpha=0.2)+
  theme_classic()
```

More detailed and additional information on colour scales can be found here: https://ggplot2-book.org/scale-colour.html
